---
title: "Building a dictionary file"
author: "Jürgen Wilbert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building a dictionary file}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
library(dplyr)
library(tibble)
library(psych)
library(scaledic)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ""
)
```

In the following examples I will lay out the construction of a dictionary file.

First, we need an example data.frame to work with (i.e. the `ex_scaledic_data` example dataset):

```{r echo=FALSE}
ex_scaledic_data %>% kable(caption = "ex_scaledic_data")
```

Second, we need a dictionary file that is also a **data.frame**. Each row of a dictionary file addresses a variable (also called *item* in this context) and each column holds a specific attribute. These attributes can be divided in two classes. Firstly, *predefined* attributes with a specific name and *additional* attributed that can have any name. The *predefined* attributes are identified by their specific names. Theses are: `item_name`, `item_label`, `values`, `value_labels`, `missing`, `weight`, and `type`[^1].

[^1]: Actually there are three more predefined attributes `class`, `scale_function`, and `scale_filter` that are not explained in this tutorial.

*Additional* attributes can have any name and are mainly used for selecting items or displaying additional information of an item (e.g., the scale and subscale an item belongs to; the reference and the author of an item; the translation of an item).

## Building a dictionary step by step

We start the example with a very simple dictionary file containing only item labels and the corresponding item names (i.e, the corresponding variable names in our data frame)[^2]. The dictionary file has two columns, `item_name` and `item_label`:

[^2]: The easiest way to create a dictionary file is to use a spreadsheet program and to import the file.

```{r echo=FALSE}
dic_file <- ex_scaledic_dic[, 1:2]
dic_file %>% kable(caption = "Dictionary file")
```

We combine the *data* file and the *dic* file with the `apply_dic()` function:

```{r}
dat_dic <- apply_dic(ex_scaledic_data, dic_file)
```

We get the message that the dic file does not contain information of the data*type* (e.g. `integer`, `factor`, `character`). The `type` attribute is used in various *scaledic* functions. For example when the data is checked for invalid values, missing values are imputed or scales are scored. A variable can be one of the following types: `integer` for numbers without decimals, `float` for numbers with decimals, `character` for variables with text, and `factor` for variables with text or numbers that are levels of a factor. 
*Scaledic* will estimate the data type from the given data when the `type` attribute is not provided in the dic file. 

Let us add `type` information to the dic file:

```{r echo=FALSE}
dic_file <- ex_scaledic_dic %>% select(1,2, type)
dic_file %>% kable(caption = "Dictionary file")
```

Most item values are of type `integer` (the item has only whole numbers) with the exception of `gender` which is of type `factor` (i.e. it has a nominal scale with several levels).

```{r}
dat_dic <- apply_dic(ex_scaledic_data, dic_file)
```

Now we can add `weight` information to the dic file. The `weight` attribute tells various *scaledic* functions a) whether the item is inverted or not, and b) how the values of an item are weighted. If a weight value is unsigned (e.g. `1`), the item is not inverted. If a weight value has a negative sign (e.g. `-1`), the item is inverted. If the (absolute) value is not `1`, the item will be weighted when calculating the item scores (e.g. `1.5` will give an item a weight of 1.5).

```{r echo=FALSE}
dic_file <- ex_scaledic_dic %>% select(1,2, weight, type)
dic_file %>% kable(caption = "Dictionary file")
```

In this example, all `weights` are `1` and no item is inverted. 

Again, we join dic and data file:

```{r}
dat_dic <- apply_dic(ex_scaledic_data, dic_file)
```

Now we add the `values` attribute to the dic file. `values` defines the valid values that a variable can take. Explicitly defining the `values` makes it possible to automatically identify invalid values in a data frame (e.g. typos). `values` are also necessary to reverse score an inverted item. For coding the `values`, you provide the possible values separated with a comma (e.g., `1, 2, 3, 4` for the integers 1 to 4). It is also possible to use a colon to define a range of integers (e.g., `5:11` indicated all integers starting with 5 and ending with 11). For variables of type `factor` or `character` you may want to provide text values. in that case, put these values within quotes (e.g., `'m', 'f', 'd'` indicates three valid text entries).

Here is the dic file with added `values`:

```{r echo=FALSE}
dic_file <- ex_scaledic_dic %>% select(1,2, "values", "weight", "type")
dic_file %>% kable(caption = "Dictionary file")
```

```{r}
dat_dic <- apply_dic(ex_scaledic_data, dic_file)
```

Invalid values can be replaced automatically with the `check_values()` function.

```{r}
dat_dic <- check_values(dat_dic, report = TRUE, replace = NA) 
```


```{r}
dat_dic %>% kable(caption = "data frame with replaced invalid values")
```

Next, we add `value_labels`. Value labels give longer labels for all or some of the values. Value labels are coded in the form `value = label` with a semicolon separating each entry. `m = male; f = female; d = diverse` codes three value label (quotes are not necessary).

Here is the dic file with added `value_labels`:

```{r echo=FALSE}
dic_file <- ex_scaledic_dic %>% select(1,2, "values", "value_labels", "weight", "type")
dic_file %>% kable(caption = "Dictionary file")
```

```{r}
dat_dic <- apply_dic(ex_scaledic_data, dic_file) %>% check_values(replace = NA)
```

Now lets see the coding for some of the variables:

```{r}
dat_dic$rel_1
```

Valid values are all integers from 1 to 6 and each value has a label.

```{r}
dat_dic$sui_1
```

Here, valid values are `0, 1, 2, 3, 4` (short `0:4`) but only the poles (`0` and `4`) have labels.

```{r}
dat_dic$gender
```

`gender` is of type factor. The valid values `'m', 'f', 'd'` have been turned into three factor levels with the corresponding labels. 

> Caution: The labels provided in the `value_labels` attribute are not automtically turned into the levels of a factor. This is because R internally would turn the values (here `'m', 'f', 'd'`) into integers (here `1, 2, 3`)

```{r}
dat_dic$age
```

`age` can take integers from 5 to 11 (`5:11`). Value labels have not been provided.

It is a common practice to code missing values with a specific number (rather than just leaving an empty entry on a datasheet). In the example dataset used here, `-999` is used as a missing value. To account for this, we add the `missing` attribute to the *dic* file containing the missing number (e.g., `-999`). Multiple missing values are separated by commas (e.g. `-99, -77`).

```{r echo=FALSE}
dic_file <- ex_scaledic_dic %>% select(1,2, values, value_labels, weight, type, missing)
dic_file %>% kable(caption = "Dictionary file")
```

The values within the `missing` attributes are automatically replaced with NA when joining a data with a dic file: 

```{r}
dat_dic <- apply_dic(ex_scaledic_data, dic_file)
```

To turn off this behavior, set the argument `replace_missing = FALSE`. Default, values are not checked for invalid entries. To do this, either set `check_values = TRUE` or apply the `check_values()` function.

```{r}
dat_dic <- apply_dic(ex_scaledic_data, dic_file, check_values = TRUE)
```

```{r}
dat_dic %>%
  slice(9:20) %>% 
  kable(caption = "Extract from the data frame with replaced missing values and checked values")
```

In the last step of this tutorial we will add information about the scales that the items belong to. Therefore, we add a new attribute called `scale` to the dic file and another attribute `scale_label` for a longer description (we could have named these attributes in any other way as they are not *predefined* attributes):

```{r echo=FALSE}
dic_file <- ex_scaledic_dic
dic_file %>% kable(caption = "Dictionary file")
```

```{r}
dat_dic <- apply_dic(ex_scaledic_data, dic_file, check_values = TRUE)
```

When can use the scale attribute to select items:

```{r}
dat_dic %>% 
  select_items(scale == "rel") %>% 
  descriptives()
```

and the `rename_items()` function to get the item labels:

```{r}
dat_dic %>% 
  select_items(scale == "rel") %>% 
  rename_items() %>%
  descriptives() %>%
  kable()
```
