---
title: "Building a dictionary file"
author: "Jürgen Wilbert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building a dictionary file}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
library(dplyr)
library(tibble)
library(psych)
library(scaledic)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What is a dictionary file?

When you do questionnaire-based research (and you are working in R), you typically create a *data.frame* with one column (variable) for each item on that questionnaire and one row for each person who participated. You can only store a limited amount of additional information about each item in that questionnaire within a data.frame (or tibble). What you can do is a) give a variable name, and b) define a variable as a factor with corresponding levels.

But basically that is it. You cannot, at least not conveniently, include a longer label for each item, the name of a scale this item belongs to, information about reverse coding, etc.

I call the collection of this additional information about items an *item dictionary*. A dictionary contains a short label, a longer description, scale affiliation, and more for each item. It is a table with one row per item and one column for each additional information about those items.   
The following table shows the information contained in a dictionary file: 

```{r tab_dic_param, echo = FALSE}
out <- tribble(
  ~Parameter, ~Meaning, ~Example,
  "item_name", "A short item label", "itrf_1",
  "scale", "Abreviation of the scale the item belongs to", "irtf",
  "subscale", "Abrevation of the sub scale", "int",
  "subscale_2", "Abrevation of the second order sub scale", "AD",
  "item_label", "Full text of the item", "Complains of headaches or stomach aches",
  "scale_label", "Name of the scale", "Integrated Teacher Report Form",
  "subscale_label", "Name of the sub scale", "internalizing problems",
  "subscale_2_label", "Name of the second order sub scale", "Anxious/Depressed",
  "values", "Valid response values in an R manner", "1:5 (for integers 1 to 5) 1,2,3 (for integers 1, 2, 3)",
  "value_labels", "Labels for each response value", "0 = not problematic; 1 = slightly problematic; 2 = problematic; 3 = strongly problematic",
  "missing", "Missing values", "-888, -999",
  "type", "Data type (factor, integer, float, real, character)", "integer",
  "weight", "Reversion of item and its weight", "1 (positive), -1 (reverse), 1.5 (positive, weights 1.5 times)",
  "source", "Reference", "Volpe et al. (2019)",
  "note", "Further notes", "Item has low descrimination"
)
kable(out, caption = "Columns of a dictionary file")

```

```{r}
n <- 20
n_var <- 10
dat <- matrix(sample(1:5, n * n_var, replace = TRUE), ncol = n_var)
dat <- as.data.frame(dat)
names(dat) <- LETTERS[1:n_var]

dat$text <- sample(c("m", "f", "ka"), n, replace = TRUE)

dic <- data.frame(
  item_name = names(dat),
  item_label = c(paste("Item", LETTERS[1:n_var]), "gender"),
  scale = c(rep("letters", n_var), "Misc"),
  scale_label = c(rep("Letter Values", n_var), "Miscellaneous"),
  values = c(rep("1:5", n_var), "'m', 'f', 'ka'"),
  value_labels = c(
    rep("1 = low; 2; 3; 4; 5 = high", n_var), 
    "m = male; f = female; ka = keine Angabe"
  ),
  type = c(rep("integer", n_var), "factor"),
  weight = 1,
  missing = c(rep("-999", n_var), "")
)

dat2 <- apply_dic(dat, dic)
dat2 <- check_values(dat2, report = TRUE)
replace_missing(dat2)

dat2[1]
dat2[[11]]

```


