---
title: "Building a dictionary file"
author: "Jürgen Wilbert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building a dictionary file}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
library(dplyr)
library(tibble)
library(psych)
library(sjPlot)
library(scaledic)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What is dictionary file

When you do questionnaire based research (and you are working in R) you usually create a *data.frame* with a column (variable) for every item of that questionnaire and a row for every person who participated. You can only store limited additional information on each item of that questionnaire within a `data.frame` (or a `tibble`). What you can do is to a) give a variable name, and b) define a variable as a factor with corresponding levels.

But basically thats it. You can not, at least not conveniently, include a longer label for each item, the name of a scale this item belogs to, information on reverse coding etc.

I call the collection of these additional information on items an *item dictionary*. A dictionary contains a short label, a longer description, scale affiliation, and more for each item. It is a table with one row per item and a column for each additonal information on those items.   
The following table shows the information included in a dictionary file: 

```{r echo = FALSE}
out <- tribble(
  ~Parameter, ~Meaning, ~Example,
  "label", "A short item label", "itrf_1",
  "scale", "Abreviation of the scale the item belongs to", "irtf",
  "subscale", "Abrevation of the sub scale", "int",
  "subscale_2", "Abrevation of the second order sub scale", "AD",
  "scale_label", "Name of the scale", "Integrated Teacher Report Form",
  "subscale_label", "Name of the sub scale", "internalizing problems",
  "subscale_2_label", "Name of the second order sub scale", "Anxious/Depressed",
  "item", "Full text of the item", "Complains of headaches or stomach aches",
  "index", "An index number", "1",
  "values", "Valid response values in an R manner", "1:5 (for integers 1 to 5) 1,2,3 (for integers 1, 2, 3)",
  "value_labels", "Labels for each response value", "0 = not problematic; 1 = slightly problematic; 2 = problematic; 3 = strongly problematic",
  "missing", "Missing values", "-888, -999",
  "type", "Data type (factor, integer, float, real, character)", "integer",
  "weight", "Reversion of item and its weight", "1 (positive), -1 (reverse), 1.5 (positive, weights 1.5 times)",
  "source", "Reference", "Volpe et al. (2019)",
  "note", "Further notes", "Item has low descrimination"
)
kable(out, caption = "Columns of a dictionary file")

```

Here is extract of a dic file from the included example:

```{r echo = FALSE}
dic_ITRF %>% slice(1:3)  %>% kable()
```
