---
title: "Introduction to scaledic"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to scaledic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
library(dplyr)
library(tibble)
library(psych)
library(scaledic)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  
)
```

# Note

This is a short description on how to use the scaledic package. The code examples in this package make use of the pipe operator `%>%` (see the `magrittr` package).  
Once you are used to the pipe operator, it makes code easier to read. Basically, it is just a shortcut. The pipe operator takes the return of the previous function and sets it as the **first** argument of the following function. An example:  

```{r eval=FALSE}
newdat <- select(dat, "iq", "age")
newdat <- describe(newdat)
newdat <- round(newdat, 2)
kable(newdat)
```

or even worth

```{r eval=FALSE}
kable(round(describe(select(dat, "iq", "age")), 2))
```

becomes:

```{r eval=FALSE}
dat %>%
  select("iq", "age") %>%
  describe() %>%
  round(2) %>%
  kable()
```

# How to use scaledic

You need two things:

1. a dataset,
2. a dictionary file.

## A dictionary file

A dictionary file is a table with a row for each variable and a column for each parameter of those variables. Most conviniently, you prepare a dictionary file in a spreadsheet program for later use with datasets.  

The following table shows the parameters and the naming of the columns of a dictionary file: 

```{r echo = FALSE}
out <- tribble(
  ~Parameter, ~Meaning, ~Example,
  "label", "A short item label", "itrf_1",
  "scale", "Abreviation of the scale the item belongs to", "irtf",
  "sub_scale", "Abrevation of the sub scale", "int",
  "sub_scale_2", "Abrevation of the second order sub scale", "shy",
  "scale_label", "Name of the scale", "Integrated Teacher Report Form",
  "sub_scale_label", "Name of the sub scale", "internalizing problems",
  "scale_2_label", "Name of the second order sub scale", "shyness",
  "item", "Full text of the item", "Vermeidet die Teilnahme an Diskussionen im Unterricht",
  "index", "An index number", "1",
  "values", "Valid response values in an R manner", "1:5 (for integers 1 to 5) 1,2,3 (for integers 1, 2, 3)",
  "value_labels", "Labels for each response value", "0 = nicht; 1 = leicht; 2 = mäßig; 3 = stark",
  "missing", "Missing values", "-888, -999",
  "type", "Data type (factor, integer, float, real)", "integer",
  "weight", "Reversion of item and its weight", "1 (positive), -1 (reverse), 1.5 (positive, weights 1.5 times)",
  "source", "Reference", "Casale et al. (2017)",
  "note", "Further notes", "Item has low descrimination"
)

kable(out, caption = "Columns of a dictionary file")

```

Here is extract of a dic file from the included example:

```{r}
dic_ITRF %>% slice(1:3)  %>% kable()
```


## Apply a dictionary file

When you combine a dataset with a dictionary file, each variable in the dataset which corresponds to a variable described in the dictionary will be complemented with the given dictionary information.  
The resulting dataset now is ready to be used with all further `scaledic` functions.  

The `apply_dic` function takes the name of the dataset and the dictionary file and combines them:

```{r}
# Here we use the example dataset "ITRF" and the example dic file "dic_ITRF"
dat <- apply_dic(ITRF, dic_ITRF)
```

Let us get an overview of all scales in the dataset:

```{r}
list_scales(dat, labels = TRUE) %>%
  kable()
```

## Clean raw data

Firstly, we check for invalid values in the dataset (e.g., typos) and replace these with NA:

```{r}
dat <- check_values(dat, replace = NA)
```

... and replace missing values as defined in the dic file with NAs

```{r}
dat <- replace_missing(dat)
```

Now we impute missing values:

```{r message=FALSE, warning=FALSE}
# Imputation for items of the subscale ITRF_Ext
dat <- impute_missing(dat, subscale = "Ext")

# Imputation for items of the subscale ITRF_Int
dat <- impute_missing(dat, subscale = "Int")

```

## Select scales for analyszing

Let us see descriptive statistics for the Internalizing sub scale:
 
```{r}
dat %>% 
  select_scale(subscale = "Int") %>%
  psych::describe(fast = TRUE) %>%
  kable()
```

## See items instead of labels

It is more convinient to see the original items instead of the short lables:

```{r}
dat %>% 
  select_scale(subscale = "Int") %>%
  names2item() %>%
  psych::describe(fast = TRUE) %>%
  kable()
```

And further we analyse the factor structure  


```{r}
dat %>%
  select_scale(scale = "ITRF") %>%
  names2item(chars = 40, short = TRUE) %>%
  psych::fa(nfactors = 2) %>%
  print(sort = TRUE, cut = 0.3)
```

## Build scale scores

Now we build scores for internalizing and externalizing scales

```{r}
dat$itrf_ext <- score_scale(dat, subscale = "Ext")
dat$itrf_int <- score_scale(dat, subscale = "Int")

```

and get descriptives for these scores

```{r}
dat %>%
  select_scores() %>%
  psych::describe(fast = TRUE) %>%
  kable()
```


